### 2016-02-01 text/MS.Rmd
version 2 of manuscript, with full summary, results, and half discussion

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
load('../output/RDatafiles/analyses.RData')
load('../output/RDatafiles/numbers.RData')
source('../R/functions-analyses.R')
source('../R/functions-figures.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=TRUE)
```
**Temperature effects on mass-scaling exponent of metabolic rates: a manipulative test**

@@

**Abstract**

Organismal metabolic rate often scales allometrically with body mass.  This pattern is observed in both interspecific and ontogenetic (intraspecific) studies, but the causal basis of variation in the allometric scaling of metabolism remains unresolved.  In the present study, we use a manipulative approach to test for an association between ontogenetic metabolic scaling and the rate and form of growth in three phyla of colonial marine organisms. Based on a Dynamic Energy Budget model that predicts the scaling of metabolic rate on the basis of energy allocation to assimilation, maintenance, growth, and maturation, we hypothesise that the scaling exponent of metabolic rate for colonial organisms should be dependent on growth rate for two-dimensional (encrusting) colonies, but not for three-dimensional (arborescent) colonies.  In support of this hypothesis, we found that the scaling exponent of metabolic rate decreased with increasing growth rate for encrusting colonies, but not for arborescent ones. These findings highlight the importance of growth for metabolic scaling during ontogeny, suggest that differences in patterns of growth might explain at least some of the documented variation in the scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.

@@

*3. Statistical analyses*  
To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data

@@

Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than the two-way interaction model (equation 3) ($\Delta$d.f. = `r deltaDF`; $\Delta$DIC = `r rounded(deltaDIC)`), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95% credible  intervals). On the other hand, *Bugula neritina* and *Hippopodina* sp. present opposing interactions (Figure 1). For example, parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp. decreases `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. On the other hand, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient for *Bugula neritina*.

**Compare model estimates with WBE and DEB predictions?**

@@

The Dynamic Energy Budget (DEB) theory predicts individual metabolism on the basis of energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`. On the basis of this, DEB theory predicts a complex three-way interaction between growth form, temperature, and size `r myCite(c('white2011amnat', 'kearney2012amnat'))`. Thus, DEB not only predicts variation in the scaling exponent of metabolic rate, it does so in a quantitative way. In support of this prediction of DEB, we show that the effect of temperature on the scaling of metabolic rate in colonial organisms depends on the growth form of the colony. The scaling exponent of metabolic rate is temperature-dependent for encrusting colonies (Figure 2A), where the mass of growing units is proportional to colony $M_i^{\sim 0.5}$, but not for arborescent colonies, where the total mass of the growing units is proportional to colony $M_i^{\sim 1}$ (Figure 2B). Such patterns cannot be predicted by models in which the scaling exponent of metabolic rate is assumed to be invariant.

We found that both $\alpha$ and $B_o$ covary with each other and with temperature in encrusting species -- this covariance has also been noted across other species and has some interesting implications for life-history theory. $B_o$ represents the mass-independent metabolic rate of an organism of unit mass, and $\alpha$ represents the power function by which metabolism increases with mass. At higher temperatures, $B_o$ increases and $\alpha$ decreases -- in other words, the metabolic costs of existing, regardless of size, increase with temperature and the metabolic efficiency benefits of increased size increase with temperature. Thus, all else being equal, increases in temperature increase the metabolic benefits of attaining a larger size while decreases in temperature reduce mass-independent metabolic costs as well as reduce the metabolic benefits of attaining a larger size. We would therefore predict that species living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction etc) while species living at lower temperatures will be under less pressure to allocate resources to size. Of course, there are many extrinsic factors that will affect selection on body size (e.g. predation, competition, etc) and a sizable theory discusses this issue but there are few formal considerations of the implications of temperature of $\alpha$ for life-history theory. Nevertheless, such differentials in the costs and benefits of attaining large size may help to explain the slow pace of life in species living in cold temperatures.

**A little example to consider.**

- Hot temperature, a = 0.6, B = 0.4. Cold temperature, a = 0.4, B = 0.6.  
- Energetic savings of being 2 units of size versus 10 units in size  
- Metabolic cost of being alive  for  2 unit size critter at high temperature is 0.79 or ~40% of total mass  
- Metabolic cost of being alive for 2 unit size critter at low temperature is 0.60 or ~30% of total mass  
- Metabolic cost of being alive for 10 unit size critter at high temperature is 1.5 or ~15% of total mass  
- Metabolic cost of being alive for 10 unit size critter at low temperature is 1.59 or ~16% of total mass  
- So 5 fold increase in size gets you a 2.667 increase in efficiency at high temperatures but same increase in size at low temperature gets you only a 1.85 increase in efficiency. That's a pretty big deal and we should say something about it!  
- Another example, at high temperatures, 10 colonies of 100 units in size will have a cumulative metabolism of 37 while 100 colonies of 10 units in size will have a cumulative metabolism of 150. (same overall mass). At low temperatures with same colony sizes and numbers, they are  63 and 159.  The ratio goes from 4.05 to 2.52. Massive change in metabolism if fragmentation occurs at high temperature, less change if fragmentation occurs at low temperatures.

New paras:  
- Many organisms exhibit multiple growth forms (encrusting and arborescent); disadvantage of arborescent is that there is no increase in efficiency with size.  
- Lack of evidence for WBE. Mass-scaling exponent not universal across the species analysed, and clear interactions with temperature.  
- Winners and losers across latitudinal gradients, and under increased temperatures. Arborsecents are predicted to be less common in warm waters,  
- Pocilliporids fast growing arborescent coral; massive corals are slow growing and something… This provides a mechanistic explanation for why arborescent corals grow fast, but attain smaller sizes than massive corals.  
- Rosa, R., and Seibel, B. A. 2009. Slow pace of life of the Antarctic colossal squid. J. Mar. Biol. Ass. U. K.

@@

*Figure 1.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*. Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively. *Top-left*: posterior distributions of mass-scaling exponents, with horizontal bars representing 95% credible intervals. *Bottom-right*: posterior distributions of metabolic normalisations, with vertical bars representing 95% credible intervals. *Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 6,000 MCMC iterations (translucent circles), with overlaid density heat-map. *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$. Regression fits were plotted using the average from the posterior distributions. In  probability posterior plots (top-left and bottom-right), smooth probability curves were approximated from the posterior density curves using very small bins (see electronic supplementary material).

>>>>

```{r, results="hide", warnings=FALSE, echo=FALSE, message=FALSE}
load('../output/RDatafiles/analyses.RData')
load('../output/RDatafiles/numbers.RData')
source('../R/functions-analyses.R')
source('../R/functions-figures.R')
cleanbib()
myRefs   <-  listBibs()
exporBibs(myRefs, erase=FALSE)
```
**Temperature effects on mass-scaling exponent of metabolic rates: a manipulative test**

@@

**Summary**

1. Organismal metabolic rate often scales allometrically with body mass, both inter and intra-specifically. Metabolic rates are also known to increase with temperature, but most theoretical tests of mass scaling and temperature dependence have treated both phenomena independently.  
2. In the present study, we use a manipulative approach to test for an association between metabolic mass scaling and temperature dependence in four species belonging to two phyla of colonial marine invertebrates.  
3. We show that mass-scaling exponents and temperature dependencies interact in different ways for different species. In so doing, we discuss the potentials of our results to multi-level scaling theory, showing that some species might be winners while others might be losers under different temperature scenarios.  
4. These findings highlight the need for new metabolic theory that formally incorporate interactions between mass-scaling and temperature dependence in ectothermic organisms, in order to better explain documented variation in the mass-scaling exponent of metabolic rate, and lead to novel and testable predictions for associations between individual metabolism and community structure.

**Key-words:** energetics, scaling, thermal limits, latitude, metabolic theory

@@

*3. Statistical analyses*  
**Things to possibly improve -- since we use a Bayesian approach, why let parameters be estimated from data (i.e. vague priors)? Why not adopt strong priors? Maybe letting it estimate is better to check whether parameters follow one or another theory?**  
**Things to possibly improve -- Include more models for model selection? E.g. two-way models with species*temp interaction or species*mass? lme4 already indicates that they are not better, I'm pretty sure the Bayesian DIC will follow results along the same lines. We'd be covering our bases better beforehand before a reviewer asks for this**

To test for the interacting effects of body mass and temperature on metabolic rate across species, we fitted a three-way interaction linear model using log-transformed data

@@

Our model selection approach indicated that the three-way interaction model (equation 2) performed significantly better than the two-way interaction model (equation 3) ($\Delta$d.f. = `r deltaDF`; $\Delta$DIC = `r rounded(deltaDIC)`), thus implying that different species present different interactions between mass-scaling and temperature dependence. Particularly, the best model yields no interaction between mass scaling and temperature dependence for *Bugula stolonifera* and the Microcionidae sponge (large overlap between 95% credible intervals for the mass-scaling slopes, $\alpha$) (Figure 1). Still, while the mass-scaling slopes are statistically the same between temperatures for these species, there is evidence for higher metabolic normalisations (i.e. intercepts), $B_o$, at higher temperatures (though not statistically significant for Microcionidae). For example, metabolic rates increase `r rounded(exp(averageEstimates$lnB25[averageEstimates$species == 'Stolonifera'])/exp(averageEstimates$lnB10[averageEstimates$species == 'Stolonifera']))`-fold between 10 ˚C and 25 ˚C for *B. stolonifera*, yielding a $Q_{10}$ of `r rounded(q10andErs$q10[q10andErs$species == 'Stolonifera'])`. It is worth noting that this $Q_{10}$ value is equivalent to an activation energy of `r rounded(q10andErs$er[q10andErs$species == 'Stolonifera'], 2)` eV (i.e. $E_r = \frac{k \times 283.15 \times 298.15 \times 1.5 \times \textrm{ln}Q_{10}}{(298.15 - 283.15)}$, where $E_r$ is an activation energy and $k$ is the Boltzmann constant, 8.62 $\times$ 10^-5^ eV K^-1^), which is consistent with the average 06 -- 0.7 eV reported in the Metabolic Theory of Ecology literature `r myCite('yvondurocher2012a')`. Different from the above species, the best model indicates that *Bugula neritina* and *Hippopodina* sp. present opposing interactions between mass-scaling and temperature dependence (Figure 1). For example, average posterior parameter estimates indicate that the difference in metabolic rates between 25 ˚C and 10 ˚C for *Hippopodina* sp. decreases, on average, `r rounded(tempDeltas[1,'tempDelta'])`-fold moving from `r round(exp(2))` to `r round(exp(5))` mg. For *Bugula neritina*, however, rates increase `r rounded(1/tempDeltas[3,'tempDelta'])`-fold along the same body-mass gradient.

Posterior distributions for mass-scaling exponents show varied consistency with predictions from the two metabolic theories. For example, *Hippopodina* sp. shows exponents that encompass both 0.67 (**DEB prediction??**; Fractal geometry model prediction for 2-D organisms; `r myCite('koontz2009amjbot')`) and 0.75 (Fractal geometry model prediction for 3-D organisms `r myCite(c('west1997science', 'west1999science'))`) at 10 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's10_97.5'], 2)`), but that are significantly shallower than 0.75 at 25 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bryo', 's25_97.5'], 2)`). On the other hand, *Bugula neritina* presents scaling exponents that overlap both 0.67 and 0.75 at 25 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's25_97.5'], 2)`), but are significantly shallower than these values at 10 ˚C (95% CI: `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Bugula', 's10_97.5'], 2)`). Microcionidae presents exponents that only overlap 0.67 at 10 ˚C, but not at 25 ˚C (95% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's10_97.5'], 2)`; 95% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Sponge', 's25_97.5'], 2)`). Finally, *Bugula stolonifera* presents values that overlap both 0.67 and 0.75 at both temperatures (95% CI at 10 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's10_97.5'], 2)`; 95% CI at 25 ˚C: `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_2.5'], 2)` -- `r rounded(ci95Estimates[ci95Estimates$species == 'Stolonifera', 's25_97.5'], 2)`).

@@

Here, we show experimentally that the mass-scaling of metabolic rates changes between two extreme temperatures, and in different ways across different species of marine invertebrates. We build on recent empirical evidence documenting similar changes (e.g. `r myCite('killen2010ele')`), yielding further support to the need for new metabolic theories (or reformulation of current ones) that formally incorporate interactive effects between mass scaling and temperature dependence of biological rates. *Currently, the Dynamic Energy Budget (DEB) theory predicts individual metabolism on the basis of energy allocation to assimilation, maintenance, growth, and maturation `r myCite('kooijman2009book')`, yielding a complex three-way interaction between growth form, temperature, and size `r myCite(c('white2011amnat', 'kearney2012amnat'))`.* **If we want to keep this comment on DEB, we will have to formally present how and what predictions DEB makes for changes in mass-scaling between different temperatures**. Our results, however, do not indicate consistent mass-temperature interactions across growth forms. Possibly, another important axis involved, which has not been considered here, is the acclimation capacity of species upon acute temperature changes (e.g. `r myCite('rezende2014functecol')`). Our findings are constrained to acute-related changes, and therefore do not consider temporal acclimation or even adaptation to changes in environmental temperature.

The fact that $B_o$ and $\alpha$ covary in different ways depending on species and temperature (Figure 1) suggests that the metabolic costs of maintenance will vary for different species at different environmental temperatures. For example, *Bugula neritina* presented a shallower mass-scaling exponent at colder temperatures, possibly indicating that large cold-environment individuals are more efficient in spending energy than counterparts in warm environments. Thus, for this species, all else being equal, increases in temperature decrease the metabolic benefits of attaining a larger size while decreases in temperature reduce the metabolic normalisation. On the other hand, our results indicate that *Hippopodina* sp. presents a reduction in metabolic benefits of attaining a larger size at colder temperatures. Thus, for this encrusting species, one could predict that individuals living at higher temperatures should allocate more resources to attaining larger sizes (as opposed to reproduction, for instance) while individuals living at lower temperatures will be under less pressure to allocate resources to size. Importantly, though, we note that there are many extrinsic factors that may affect selection on body size (e.g. predation, competition, etc). Yet, such differentials in the costs and benefits of attaining large size may help explain why in species living in cold temperatures present a slower pace of life (e.g. `r myCite('rosa2010jmba')`). Moreover, they may represent necessary improvements on current attempts to scale up energy from individuals to higher levels of organisation (e.g. `r myCite(c('allen2005','yvondurocher2012b','barneche2014ele'))`).

Importantly, some of our results are also consistent with existing theoretical predictions. 
- Temperature dependence and mass-scaling of *B. stolonifera* consistent with MTE. Why this species only?

Conclusions paragraph

@@

*Figure 1.* Estimates from the three-way interaction model as obtained by Bayesian methods in *JAGS*. Results for each species are represented in groups of four plots, with blue and red colours representing actual data and parameter estimates at 10 ˚C and 25 ˚C, respectively. *Top-left* and *bottom-right*: posterior probabilities of mass-scaling exponents and metabolic normalisations (respectively), with bars over curves representing 95% credible intervals, and numbers representing average estimates. *Bottom-left*: covariance between mass-scaling exponents and metabolic normalisations for each of the 6,000 MCMC iterations (translucent circles), with overlaid density heat-map. *Top-right*: regressions of individual metabolic rate, $\textrm{ln}B_i$, as functions of individual body mass, $\textrm{ln}M_i$. Regression fits were plotted using the average from the posterior distributions. In probability posterior plots (*top-left* and *bottom-right*), smooth probability curves (*n* = 511 values) were approximated from posterior density curves (*n* = 512 values); differences between adjacent $x$ values were multiplied by the average of adjacent $y$ values (i.e. $(x_{i+1} - x_{i}) \times ((y_{i+1} - y_{i})/2)$) thus insuring that the integral of the probability curve $\approx$ 1.
